Microbiome analysis - Mirzapur follow-up

### 2020-09-14 code/R_markdown/P3_machine_learning_figures.Rmd
updates re focusing on top 15 imp var and adding crypto species to rf analysis

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_plot_2020_09_02.RData")

@@

plot_tax_4D %>%
  ggplot() +geom_bar(aes(x=var, y = value, fill = variable),stat = "identity", width=0.4, position = position_dodge(width=0.5)) +
  scale_fill_manual(values = c("red","blue")) +
  xlab(NULL) + ylab('Mean decrease in node impurity') + #facet_wrap(~tod_or_pre, scales = "free_x") +
  coord_flip() + 
  theme(legend.position = c(.76, 0.13), 
        axis.text.x = element_text(size = 7),
        legend.title=element_text(size=10), 
        legend.text=element_text(size=8))  + scale_x_discrete(limits = rev(levels(plot_tax_4D$var)))
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig4D_imp_var_microbiome_mirpur.png",
       width = 4.5,height = 4,units = "in",dpi = 300)

@@

ggplot(twoASVs, aes(x = var, y = value_use, color = symptomatic_use)) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_point(alpha = 0.5, position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0, dodge.width = 0.75, seed = SEED_INT)) + 
  facet_wrap(~time) +
  scale_color_manual(values= colors[1:2]) +
  scale_y_sqrt() + ylab('relative abundance') + xlab(NULL) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        legend.position = "bottom",
        legend.title = element_text(size = 9)) + guides(color=guide_legend(title="symptomatic?")) +
  geom_hline(yintercept = 0.2, linetype = 'dotted', alpha = .75)
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig4E_mega.png",
       width = 3.2,height = 4,units = "in",dpi = 300)

ggplot(two_WIDE, aes(x = `time-of-detection`, y = `pre-detection`, color = `symptomatic?`)) +
  geom_jitter(alpha = 0.5, position = position_jitter(width = 0.01, height = 0.01, seed = SEED_INT)) +
  scale_color_manual(values = colors[1:2]) + scale_x_sqrt() + scale_y_sqrt()+
  theme(panel.grid.minor = element_blank(),
        legend.position="bottom")

ggplot(two_WIDE, aes(x = `time-of-detection`, y = `pre-detection`, color = `symptomatic?`)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = colors[1:2]) + scale_x_sqrt() + scale_y_sqrt()+
  theme(panel.grid.minor = element_blank(),
        legend.position="bottom")

@@

load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_TOD_abx_2020_09_02.RData")
AUC_TOD_abx = rf[[3]]
rm(rf)
roc_list = list(roc_TOD,roc_TOD_abx)
names(roc_list) = c("TOD microbiota","TOD microbiota &\nantibiotic history")
p = pROC::ggroc(roc_list, aes = "colour") + 
  theme_bw() +
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), color="grey", linetype="dashed") +
  scale_color_manual(values = c("blue","grey")) + 
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm"),
        legend.title = element_blank()) +
  annotate("text", label = paste0("AUC = ",round(AUC_TOD, digits = 2)), x = .2, y = .27, color = "blue") +
  annotate("text", label = paste0("AUC = ",round(AUC_TOD_abx, digits = 2)), x = .2, y = .2, color = "grey") 
p
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/SFig5A_ROC_TOD_abx_mirpur.png",
       p,width = 5,height = 3,units = "in",dpi = 300)

tax2_abx %>%
  ggplot() +geom_bar(aes(x=reorder(var,value), y = value, fill = model),stat = "identity", width=0.4, position = position_dodge(width=0.5)) +
  scale_fill_manual(values = c("black","grey")) +
  xlab(NULL) + ylab('Mean decrease in node impurity') + #facet_wrap(~tod_or_pre, scales = "free_x") +
  coord_flip()
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/SFig5B_imp_var_abx_mirpur.png",
       width = 7,height = 3,units = "in",dpi = 300)

>>>>

# load data
#load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_plot_2020_09_02.RData")
#load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_plot_2020_09_10.RData")
#load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/cmel_P3output_plot_2020_09_11.RData")
#load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/chom_P3output_plot_2020_09_11.RData")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_plot_2020_09_11.RData")
#load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/WOC_P3output_plot_2020_09_11.RData")

@@

plot_tax_4D %>%
  ggplot() +geom_bar(aes(x=var, y = value, fill = variable),stat = "identity", width=0.4, position = position_dodge(width=0.5)) +
  scale_fill_manual(values = c("red","blue")) +
  xlab(NULL) + ylab('Mean decrease in node impurity') + #facet_wrap(~tod_or_pre, scales = "free_x") +
  coord_flip() + 
  theme(legend.position = c(.76, 0.13), 
        axis.text.x = element_text(size = 7),
        legend.title=element_blank(), 
        legend.text=element_text(size=8))  + scale_x_discrete(limits = rev(levels(plot_tax_4D$var)))
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig4D_imp_var_microbiome_mirpur.png",
       width = 4.5,height = 4,units = "in",dpi = 300)

@@

ggplot(twoASVs, aes(x = var, y = value_use, color = symptomatic_use)) + 
  geom_boxplot(outlier.shape = NA) + 
  geom_point(alpha = 0.5, position = position_jitterdodge(jitter.width = 0.2, jitter.height = 0, dodge.width = 0.75, seed = SEED_INT)) + 
  facet_wrap(~time, scales = "free_x") +
  scale_color_manual(values= colors[1:2]) +
  scale_y_sqrt() + ylab('relative abundance') + xlab(NULL) + 
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        legend.position = "bottom",
        legend.title = element_text(size = 9)) + guides(color=guide_legend(title="symptomatic?")) +
  geom_hline(yintercept = 0.2, linetype = 'dotted', alpha = .75)
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/Fig4E_mega.png",
       width = 3.2,height = 4,units = "in",dpi = 300)

@@

load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/rf_TOD_abx_2020_09_11.RData")
#load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/WOC_rf_TOD_abx_2020_09_11.RData")

AUC_TOD_abx = rf[[3]]
rm(rf)
roc_list = list(roc_TOD,roc_TOD_abx)
names(roc_list) = c("TOD microbiota","TOD microbiota &\nantibiotic history")
p = pROC::ggroc(roc_list, aes = "colour") + 
  theme_bw() +
  geom_segment(aes(x = 1, xend = 0, y = 0, yend = 1), color="grey", linetype="dashed") +
  scale_color_manual(values = c("blue","grey")) + 
  theme(plot.margin=unit(c(.5,.5,.5,.5),"cm"),
        legend.title = element_blank()) +
  annotate("text", label = paste0("AUC = ",round(AUC_TOD, digits = 2)), x = .2, y = .27, color = "blue") +
  annotate("text", label = paste0("AUC = ",round(AUC_TOD_abx, digits = 2)), x = .2, y = .2, color = "grey") 
p
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/SFig5A_ROC_TOD_abx_mirpur.png",
       p,width = 5,height = 3,units = "in",dpi = 300)

tax2_abx %>%
  ggplot() +geom_bar(aes(x=reorder(var,value), y = value, fill = model),stat = "identity", width=0.4, position = position_dodge(width=0.5)) +
  scale_fill_manual(values = c("grey","blue")) +
  xlab(NULL) + ylab('Mean decrease in node impurity') + #facet_wrap(~tod_or_pre, scales = "free_x") +
  coord_flip()
ggsave("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/figures/SFig5B_imp_var_abx_mirpur.png",
       width = 7,height = 3,units = "in",dpi = 300)

### 2020-09-14 code/R_markdown/P4_univariate_stats.Rmd
updates re focusing on top 15 imp var and adding crypto species to rf analysis

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P1output_2020_09_02.RData")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_p1_2020_09_02.RData")

>>>>

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P1output_2020_09_11.RData")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_p1_2020_09_11.RData")

### 2020-09-14 code/R_markdown/P4_univariate_stats_figures.Rmd
updates re focusing on top 15 imp var and adding crypto species to rf analysis

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P4output_plot_2020_09_02.RData")

@@

>>>>

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P4output_plot_2020_09_11.RData")

@@

Megasphaera fold change
```{r}

tod_tax2 %>% filter(Genus == 'Megasphaera', size2 == 1) %>% select(log2FoldChange,padj)

### 2020-09-14 code/R_markdown/P5_mirzapur_analysis.Rmd
updates re focusing on top 15 imp var and adding crypto species to rf analysis

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/data_cannot_be_shared/data.RData")
rm(list= ls()[! (ls() %in% c('left_study_ids'))])
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_p2_2020_09_02.RData")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_p3_2020_09_02.RData")
ps_mirzapur = readRDS("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/ps_2020_08_12.rds")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_p1_2020_09_02.RData") # load save_df1 and save_df2

@@

# number of Megaspaheara ASVs
temp = as.data.frame(tax_table(ps_mirzapur_rare))
dim(temp[temp$Genus == 'Megasphaera',])
# temp2 = as.data.frame(tax_table(ps_rare))
# dim(temp2[temp2$Genus == 'Megasphaera',])

@@

# get TOD values
plt_df = melt(otu_for_comp_TOD %>% rownames_to_column('sample')) %>%
  group_by(variable) %>%
  mutate(value_use = value/max(value)) %>%
  ungroup() 
tax = tax_table(ps_mirzapur_rare)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tax %>% 
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))
twoASVs_TOD = tax2 %>% filter(var == 'ASV10 (Megasphaera)')

# get PD values
plt_df = melt(otu_for_comp_PD %>% rownames_to_column('sample')) %>%
  group_by(variable) %>%
  mutate(value_use = value/max(value)) %>%
  ungroup() 
tax = tax_table(ps_mirzapur_rare)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tax %>% 
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))
PD_twoASVs = tax2 %>% filter(var == 'ASV10 (Megasphaera)')

@@

df_mirz = melt(rbind(count_df_PD_mirzapur, count_df_TOD_mirzapur))  %>% mutate(Site = 'Mirzapur')
df_mir = melt(rbind(count_df_PD_mirpur %>% select(-symptomatic_use,-Collinsella,-`symptomatic?`) %>% 
                      mutate(time = 'pre-detection'), 
                    count_df_TOD_mirpur %>% select(-symptomatic_use,-Parabacteroides,-`symptomatic?`) %>% 
                      mutate(time = 'time-of-detection'))) %>% mutate(Site = 'Mirpur')
df = rbind(df_mirz,df_mir)

>>>>

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/data_cannot_be_shared/data.RData")
rm(list= ls()[! (ls() %in% c('left_study_ids'))])
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_p2_2020_09_11.RData")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_p3_2020_09_11.RData")
ps_mirzapur = readRDS("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/ps_2020_08_12.rds")
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P3output_p1_2020_09_11.RData") # load save_df1 and save_df2

@@

ntaxa(subset_taxa(ps_mirzapur_rare, Genus == "Megasphaera"))

@@

# get TOD values
plt_df = melt(otu_for_comp_TOD %>% rownames_to_column('sample')) %>%
  group_by(variable) %>%
  mutate(value_use = value/max(value)) %>%
  ungroup() 
tax = tax_table(ps_mirzapur_rare)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tax %>% 
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))
twoASVs_TOD = tax2 %>% filter(var == tax2$var[grep("Megasphaera", tax2$var)])

# get PD values
plt_df = melt(otu_for_comp_PD %>% rownames_to_column('sample')) %>%
  group_by(variable) %>%
  mutate(value_use = value/max(value)) %>%
  ungroup() 
tax = tax_table(ps_mirzapur_rare)
tax = tax[rownames(tax) %in% plt_df$variable,]
tax = merge(tax,plt_df, by.x= 'row.names', by.y = "variable", all.x = F, all.y = T)
tax$Genus = as.character(tax$Genus)
tax2 = tax %>% 
  transform(id=as.numeric(factor(Row.names))) %>%
  mutate(var = ifelse(is.na(Family),
                      as.character(Row.names),
                      ifelse(is.na(Genus), paste0(paste0(paste0(paste0(as.character('ASV'),id),' ([f]'),Family),')'),
                             paste0(paste0(paste0(paste0(as.character('ASV'),id),' ('),Genus),')'))))
PD_twoASVs = tax2 %>% filter(var == tax2$var[grep("Megasphaera", tax2$var)])

@@

#UPDATE HERE: count_df_PD_mirpur column names have switched from genus to ASV
count_df_PD_mirzapur
df_mirz = melt(rbind(count_df_PD_mirzapur, count_df_TOD_mirzapur))  %>% mutate(Site = 'Mirzapur')
df_mir = melt(rbind(count_df_PD_mirpur %>% 
                      rename(Megasphaera= `ASV15 (Megasphaera)`) %>%
                      select(sample, Megasphaera) %>% 
                      mutate(time = 'pre-detection'), 
                    count_df_TOD_mirpur %>% select(-symptomatic_use,-Parabacteroides,-`symptomatic?`) %>% 
                      mutate(time = 'time-of-detection'))) %>% mutate(Site = 'Mirpur')
df = rbind(df_mirz,df_mir)

### 2020-09-14 code/R_markdown/P5_mirzapur_analysis_figures.Rmd
updates re focusing on top 15 imp var and adding crypto species to rf analysis

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P5output_plot_2020_09_02.RData")

@@

>>>>

# load data
load("/Users/maureencarey/local_documents/work/crypto_microbiome/cryptosporidium_microbiome/results/P5output_plot_2020_09_11.RData")

@@

ntaxa(subset_taxa(ps_mirzapur_rare, Genus == "Megasphaera"))
